Binocular integration of chromatic and luminance signals

### 2024-03-10 manuscript.Rmd
Various updates and improvements

---
title: "Binocular integration of chromatic and luminance signals"
author: |
   | Daniel H. Baker, Kirralise J. Hansford, Federico G. Segala, Annie Y. Morsi,
   | Rowan J. Huxley, Joel T. Martin, Maya Rockman & Alex R. Wade
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    fig_caption: yes
    toc: no
    keep_tex: yes
  word_document: default
  html_document: default
bibliography: references.bib
csl: journal-of-vision.csl
header-includes: 
  \usepackage{float} \floatplacement{figure}{H} 
  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{A\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{A\arabic{figure}}}
---

@@

processdata <- 0
# the processdata flag has three settings:
# 0: do no analysis, just produce the pdf of the manuscript
# 1: fit psychometric functions and plot graphs
# 2: do Stan modelling (takes N hours)

nsamples <- 10000    # number of Stan samples for modelling

# reasonably compact code to check which packages are installed, install the missing ones, and activate all
packagelist <- c('knitr','reticulate','osfr','bookdown','rstan','coda','modeest','pals','pracma','doParallel','R.matlab') # list of CRAN packages
missingpackages <- packagelist[!packagelist %in% installed.packages()[,1]]
if (length(missingpackages)>0){install.packages(missingpackages)}
toinstall <- packagelist[which(!packagelist %in% (.packages()))]
invisible(lapply(toinstall,library,character.only=TRUE))

@@

@@

allthresh[subj,expt,cond,pedlevel] <- res$Fit[1]
allslope[subj,expt,cond,pedlevel] <- res$Fit[2]/C
}

@@

meanthresh <- apply(allthresh,2:4,mean)
meanslope <- apply(allslope,2:4,mean)

@@

  MCSthresh[subj,cond] <- res$Fit[1]
  MCSslope[subj,cond] <- res$Fit[2]/C

@@

slopematrix[subj,1,] <- 10.3/MCSslope[subj,c(4,7,10)]
slopematrix[subj,2,] <- 10.3/MCSslope[subj,c(5,8,11)]
slopematrix[subj,3,] <- 10.3/MCSslope[subj,c(6,9,12)]
}
detslopes <- 10.3/MCSslope[,1:3]

save(file='local/MCSdata.RData',list=c('MCSthresh','MCSslope','thdelev','slopematrix','detslopes','MCSpropcorr','contrastsdB'))

@@

  modelresp <- -10
  contrastinc <- 0
  if (baseline>-999){

    while (((modelresp-baseline) < p[7])){
      contrastinc <- contrastinc + 0.1
  if (cond==1){modelresp <- getmodelresp(p,pedC+contrastinc,0)}
  if (cond==2){modelresp <- getmodelresp(p,pedC+contrastinc,pedC+contrastinc)}
  if (cond==3){modelresp <- getmodelresp(p,pedC+contrastinc,pedC)}
  if (cond==4){modelresp <- getmodelresp(p,pedC,contrastinc)}
    }
    while (((modelresp-baseline) > p[7])){
      contrastinc <- contrastinc - 0.001
  if (cond==1){modelresp <- getmodelresp(p,pedC+contrastinc,0)}
  if (cond==2){modelresp <- getmodelresp(p,pedC+contrastinc,pedC+contrastinc)}
  if (cond==3){modelresp <- getmodelresp(p,pedC+contrastinc,pedC)}
  if (cond==4){modelresp <- getmodelresp(p,pedC,contrastinc)}
    }    
    }
  else{contrastinc <- 99}
return(contrastinc) }

@@

ntotalsimplexfits <- 100

cluster <- makeCluster(8)
registerDoParallel(cluster)

@@

par(pty="s",mfcol=c(1,3))  # make axis square
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(-12,36), ylim=c(-12,39))   
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.3)     
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)
title(xlab="Pedestal contrast (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)   
title(ylab="Threshold (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)
title(main='Achromatic')

@@

par(pty="s")  # make axis square
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(-12,36), ylim=c(-12,39))  
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.3)    
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)
title(xlab="Pedestal contrast (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)   
title(ylab="Threshold (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)
title(main='Red/Green')

@@

par(pty="s")  # make axis square
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(-12,36), ylim=c(-12,39))   
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.3)     
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)
title(xlab="Pedestal contrast (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)
title(ylab="Threshold (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)
title(main='Blue/Yellow')

@@

  numSavedSteps <- nsamples    # total number of MCMC steps
  adaptSteps = 200  # Number of steps to "tune" the samplers
  burnInSteps = 200 
  thinSteps <- 1
  nChains = 24
  options(mc.cores=8)

modelString = "
  data {
    int<lower=1> Nsubj ;
    int<lower=1> Ntotal ;
    int correct[Ntotal] ;
    real PedestalContrast[Ntotal] ;
    real TargetContrast[Ntotal] ;
    int<lower=1> c[Ntotal] ;
    int<lower=1> s[Ntotal] ;
  }
  parameters {
    real<lower=0> p[Nsubj] ;
    real<lower=0> q[Nsubj] ;    
    real<lower=0> m[Nsubj] ;
    real<lower=0> S[Nsubj] ;  
    real<lower=0> Z[Nsubj] ;
    real<lower=0> w[Nsubj] ;
    real<lower=0> k[Nsubj] ;

@@

    real Lt ;
    real Rt ;
    real Ln ;
    real Rn ;
    real targresp ;
    real nullresp ;

@@

    targresp = pow((Lt + Rt), (p[s[i]]+q[s[i]])) / (Z[s[i]] + pow((Lt + Rt), q[s[i]])) ;
    nullresp = pow((Ln + Rn), (p[s[i]]+q[s[i]])) / (Z[s[i]] + pow((Ln + Rn), q[s[i]])) ;     
    correct[i] ~ bernoulli((1 + erf((targresp - nullresp)/k[s[i]]))/2) ;

@@

  ex1data <- read.csv('local/Experiment1.csv')
  explist <- unique(ex1data$Experiment)

for (mod in 1:1){
  expdata <- subset(ex1data,Experiment==explist[mod])
  expdata$Participant <- as.numeric(gsub('\\D','',expdata$Subject))
  expdata$Condition <- ceiling(expdata$Condition/2)

  dataList = list(correct = expdata$IsCorrect, PedestalContrast = expdata$PedestalContrast, TargetContrast = expdata$TargetContrast, c = expdata$Condition, s = expdata$Participant, Nsubj = max(expdata$Participant), Ntotal = nrow(expdata))

  # Get MC sample of posterior:
    stanFit <- sampling(object=dipmodel, 
                        data = dataList, 
                        chains = nChains,
                        iter = (ceiling(numSavedSteps/nChains)*thinSteps
                                +burnInSteps), 
                        warmup = burnInSteps, 
                        thin = thinSteps,
                        cores = getOption("mc.cores", 1L))

    # convert stan format to coda format:jkn9
    mcmcCoda2 = mcmc.list(lapply( 1:ncol(stanFit), function(x){mcmc(as.array(stanFit)[,x,])}))

    save(file=paste0('local/E1mod',mod,'.RData'),list=c('stanFit','mcmcCoda2'))

@@

```{r plotdata, include=FALSE}

@@

  paramorder <- c('p','q','m','S','Z','w','k')
  allsamples <- array(0,dim=c(3,7,10008))
  for (expt in 1:3){
    load(file=paste0('local/E1mod',expt,'.RData'))
    p <- NULL
    q <- NULL
    m <- NULL
    S <- NULL
    Z <- NULL
    w <- NULL
    k <- NULL
    for (i in 1:length(mcmcCoda2)){
      p <- c(p,mcmcCoda2[[i]][,23])  
      q <- c(q,mcmcCoda2[[i]][,24])  
      m <- c(m,mcmcCoda2[[i]][,25])  
      S <- c(S,mcmcCoda2[[i]][,26])  
      Z <- c(Z,mcmcCoda2[[i]][,27])  
      w <- c(w,mcmcCoda2[[i]][,28])  
      k <- c(k,mcmcCoda2[[i]][,29])  
    }        
    p <- p + q
allsamples[expt,1,] <- p
allsamples[expt,2,] <- q
allsamples[expt,3,] <- m
allsamples[expt,4,] <- S
allsamples[expt,5,] <- Z
allsamples[expt,6,] <- w
allsamples[expt,7,] <- k

  }
# meanparams <- 10^apply(log10(allsamples),1:2,mlv,method='meanshift')
meanparams <- 10^apply(log10(allsamples),1:2,mean)
# meanparams[2:3,5] <- meanparams[2:3,5]*100

pdf('Figures/modelparams.pdf',width=12,height=8)

@@

par(pty="s",mfcol=c(2,4))

ticklocsx <- seq(-18,18,6)
ticklabelsx <- c('1/8','1/4','1/2',1,2,4,8)
ticklocsy <- seq(0,1,0.2)
ticklabelsy <- ticklocsy

@@

  plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(-18,18), ylim=c(0,1))  
  axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.3)    
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)
title(xlab=paramorder[param], col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)

  for (expt in 1:3){

    svals <- 20*log10(allsamples[expt,param,])
    a <- density(svals)
    a$y <- a$y/max(a$y)
    lines(a$x,a$y,col=collist[expt])

    lines(20*log10(meanparams[expt,c(param,param)]),c(0,1),col=collist[expt])
  }

@@

for (expt in 1:3){
  for (cond in 1:4){
    for (pedlev in 1:length(pedlist)){
      allpred[expt,cond,pedlev] <- discriminate(meanparams[expt,],pedlevelsC[pedlev],cond)
    }
  }
}
allpreddB <- 20*log10(allpred)

pdf('Figures/dippers.pdf',width=12,height=5)

@@

par(pty="s",mfcol=c(1,3))  # make axis square
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(-12,36), ylim=c(-12,39))   
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.3)     
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)
title(xlab="Pedestal contrast (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)   
title(ylab="Threshold (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)
title(main='Achromatic')

@@

for (cond in 1:4){
  arrows(pedcontrasts,meanthresh[1,cond,],pedcontrasts,meanthresh[1,cond,]+SEthresh[1,cond,],col=collist[cond],angle=90,length=0.02,lwd=2)
 arrows(pedcontrasts,meanthresh[1,cond,],pedcontrasts,meanthresh[1,cond,]-SEthresh[1,cond,],col=collist[cond],angle=90,length=0.02,lwd=2)  
  # lines(pedcontrasts,meanthresh[1,cond,],col=collist[cond],lwd=2)
  lines(pedlist,allpreddB[1,cond,],col=collist[cond],lwd=2)
  points(pedcontrasts,meanthresh[1,cond,],pch=16,col=collist[cond],cex=1.5)
}
legend(-12,36,c('Monocular','Binocular','Half-bin','Dichoptic'),pch=16,col=collist,box.lwd=2,cex=1.5)

par(pty="s")  # make axis square
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(-12,36), ylim=c(-12,39))  
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.3)    
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)
title(xlab="Pedestal contrast (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)   
title(ylab="Threshold (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)
title(main='Red/Green')

for (cond in 1:4){
  arrows(pedcontrasts,meanthresh[3,cond,],pedcontrasts,meanthresh[3,cond,]+SEthresh[3,cond,],col=collist[cond],angle=90,length=0.02,lwd=2)
 arrows(pedcontrasts,meanthresh[3,cond,],pedcontrasts,meanthresh[3,cond,]-SEthresh[3,cond,],col=collist[cond],angle=90,length=0.02,lwd=2)  
 # lines(pedcontrasts,meanthresh[3,cond,],col=collist[cond],lwd=2)
  lines(pedlist,allpreddB[3,cond,],col=collist[cond],lwd=2)
  points(pedcontrasts,meanthresh[3,cond,],pch=16,col=collist[cond],cex=1.5)
}

par(pty="s")  # make axis square
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(-12,36), ylim=c(-12,39))   
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.3)     
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)
title(xlab="Pedestal contrast (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)
title(ylab="Threshold (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)
title(main='Blue/Yellow')

for (cond in 1:4){
  arrows(pedcontrasts,meanthresh[2,cond,],pedcontrasts,meanthresh[2,cond,]+SEthresh[2,cond,],col=collist[cond],angle=90,length=0.02,lwd=2)
 arrows(pedcontrasts,meanthresh[2,cond,],pedcontrasts,meanthresh[2,cond,]-SEthresh[2,cond,],col=collist[cond],angle=90,length=0.02,lwd=2)  
 # lines(pedcontrasts,meanthresh[2,cond,],col=collist[cond],lwd=2)
  lines(pedlist,allpreddB[2,cond,],col=collist[cond],lwd=2)
  points(pedcontrasts,meanthresh[2,cond,],pch=16,col=collist[cond],cex=1.5)
}

@@

@@

  load('local/MCSdata.RData')
slopematrix <- 10^(apply(log10(slopematrix),c(2,3),mean))
slopematrix <- round(slopematrix,digits=2)

@@

titlelist <- c('(a) Achromatic target','(b) R/G target','(c) B/Y target','(d) Threshold elevation','(e) Psychometric slope')

    pdf(paste0('Figures/MCSdata.pdf'),width=12,height=8)

    layout(matrix(c(1,1,2,2,3,3,0,4,4,5,5,0),2,6,byrow=TRUE))

@@

@@

load('local/MCSdata.RData')
slopematrix <- 10^(apply(log10(slopematrix),c(2,3),mean))
slopematrix <- round(slopematrix,digits=2)
contrastsdB <- contrastsdB[1,,]

@@

```{r dipperfig, fig.cap="Dipper functions from Experiment 1, averaged across three participants. Error bars show ±1SE across participants. Note that contrast values are expressed as a percentage of the maximum displayable contrast (see Procedures for details).", fig.align="center", echo=FALSE}

>>>>

---
title: "Binocular integration of chromatic and luminance signals"
author: |
   | Daniel H. Baker, Kirralise J. Hansford, Federico G. Segala, Annie Y. Morsi,
   | Rowan J. Huxley, Joel T. Martin, Maya Rockman & Alex R. Wade
date: "Department of Psychology, University of York"
output:
  bookdown::pdf_document2:
    fig_caption: yes
    toc: no
    keep_tex: yes
  word_document: default
  html_document: default
bibliography: references.bib
csl: journal-of-vision.csl
header-includes: 
  \usepackage{float} \floatplacement{figure}{H} 
  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{A\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{A\arabic{figure}}}
---

@@

processdata <- 2
# the processdata flag has three settings:
# 0: do no analysis, just produce the pdf of the manuscript
# 1: fit psychometric functions and plot graphs
# 2: do Stan modelling (takes N hours)

nsamples <- 100000    # number of Stan samples for modelling
ntotalsimplexfits <- 100    # number of random starting vectors for simplex fits
ncores <- 8    # number of processor cores to use on the host machine
postsampleres <- 400    # resolution of psignifit posterior sampling

# other Stan settings
adaptSteps = 5000  # Number of steps to "tune" the samplers
burnInSteps = 5000
thinSteps <- 10
nChains = 64

# reasonably compact code to check which packages are installed, install the missing ones, and activate all
packagelist <- c('knitr','reticulate','osfr','bookdown','rstan','coda','modeest','pals','pracma','doParallel','R.matlab','bayestestR','alphahull') # list of CRAN packages
missingpackages <- packagelist[!packagelist %in% installed.packages()[,1]]
if (length(missingpackages)>0){install.packages(missingpackages)}
toinstall <- packagelist[which(!packagelist %in% (.packages()))]
invisible(lapply(toinstall,library,character.only=TRUE))

@@

extract_polygons <- function(alpha_obj){
  if(class(alpha_obj) != "ashape") stop("extract_polygons requires an ashape")
  edge.df <- as.data.frame(alpha_obj$edges)
  groups <- ns <- xs <- ys <- numeric(nrow(edge.df))
  m <- cbind(edge.df[[1]], edge.df[[2]])
  group <- 1
  repeat {
    i <- which(groups == 0)[1]
    if (length(i) == 0 | is.na(i)) break()
    j <- n <- 1
    repeat {
      groups[i] <- group
      ns[i] <- n
      if(j == 1) xs[i] <- edge.df$x1[i] else xs[i] <- edge.df$x2[i]
      if(j == 1) ys[i] <- edge.df$y1[i] else ys[i] <- edge.df$y2[i]
      next_ind <- which((m[, j] == m[i, j] | m[, (j %% 2 + 1)] == m[i, j]) & groups == 0)
      if (length(next_ind) == 0) break()
      j <- which(m[next_ind,] == m[i, j]) %% 2 + 1
      i <- next_ind
      n <- n + 1
    }
    group <- group + 1
  }
  data.frame(x = xs, y = ys, group = as.factor(groups))[order(groups, ns), ]
}

@@

allthresh[subj,expt,cond,pedlevel] <- res$Fit[1]
allslope[subj,expt,cond,pedlevel] <- 10.3/(res$Fit[2]/C)
}

@@

meanthresh <- apply(allthresh,2:4,mean)
meanslope <- 10^apply(log10(allslope),2:4,mean)

@@

  MCSthresh[subj,cond] <- res$Fit[1]
  MCSslope[subj,cond] <- 10.3/(res$Fit[2]/C)

@@

slopematrix[subj,1,] <- MCSslope[subj,c(4,7,10)]
slopematrix[subj,2,] <- MCSslope[subj,c(5,8,11)]
slopematrix[subj,3,] <- MCSslope[subj,c(6,9,12)]
}
detslopes <- MCSslope[,1:3]

options$stepN <- as.integer(c(postsampleres,postsampleres,20,20,20))
allposteriors <- array(0,dim=c(12,postsampleres,postsampleres))
allposX <- array(0,dim=c(12,postsampleres))
allposY <- array(0,dim=c(12,postsampleres))

# now also pool across subjects and fit for plotting the full posterior
for (cond in 1:12){
  datatofit <- data.frame(contrastsdB[1,cond,],colSums(ncorrect[,cond,]),colSums(ntrials[,cond,]))
  colnames(datatofit) <- c('CdB','Ncorrect','Ntrials')
  res <- ps$psignifit(as.matrix(datatofit),options)

  xycos <- res$X1D
  a <- res$Posterior
  sumpos <- apply(a,c(1,2),sum)

  allposteriors[cond,,] <- sumpos
  allposX[cond,] <- xycos[[1]]
  allposY[cond,] <- xycos[[2]]

save(file='local/MCSdata.RData',list=c('MCSthresh','MCSslope','thdelev','slopematrix','detslopes','MCSpropcorr','contrastsdB','allposteriors','allposX','allposY'))

@@

@@

cluster <- makeCluster(ncores)
registerDoParallel(cluster)

@@

par(pty="s",mfcol=c(1,3))  # make axis square
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(-12,36), ylim=c(-12,39))   
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.3)     
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)
title(xlab="Pedestal contrast (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)   
title(ylab="Threshold (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)
title(main='(a) Achromatic',cex.main=2)

@@

par(pty="s")  # make axis square
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(-12,36), ylim=c(-12,39))  
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.3)    
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)
title(xlab="Pedestal contrast (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)   
title(ylab="Threshold (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)
title(main='(b) Red/Green',cex.main=2)

@@

par(pty="s")  # make axis square
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(-12,36), ylim=c(-12,39))   
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.3)     
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)
title(xlab="Pedestal contrast (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)
title(ylab="Threshold (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)
title(main='(c) Blue/Yellow',cex.main=2)

@@

numSavedSteps <- nsamples    # total number of MCMC steps
options(mc.cores=ncores)

modelString = "
  data {
    int<lower=1> Ndatasets ;
    int<lower=1> Ntotal ;
    int ncorrect[Ntotal] ;
    int ntrials[Ntotal] ;
    real PedestalContrast[Ntotal] ;
    real TargetContrast[Ntotal] ;
    int<lower=1> c[Ntotal] ;
    int<lower=1> s[Ntotal] ;    
  }
  parameters {
    real<lower=0> p[Ndatasets];
    real<lower=0> q[Ndatasets];    
    real<lower=0> m[Ndatasets];
    real<lower=0> S[Ndatasets];  
    real<lower=0> Z[Ndatasets];
    real<lower=0> w[Ndatasets];
    real<lower=0> k[Ndatasets];

@@

    real Lt ;
    real Rt ;
    real Ln ;
    real Rn ;
    real targresp ;
    real nullresp ;
    real dprime ;

@@

    targresp = pow((Lt + Rt), (p[s[i]]+q[s[i]])) / (Z[s[i]] + pow((Lt + Rt), q[s[i]])) ;
    nullresp = pow((Ln + Rn), (p[s[i]]+q[s[i]])) / (Z[s[i]] + pow((Ln + Rn), q[s[i]])) ;

    dprime = (targresp-nullresp)/k[s[i]] ;

    ncorrect[i] ~ binomial(ntrials[i], Phi(dprime/sqrt(2))) ;

@@

ex1data <- read.csv('local/Experiment1.csv')
ex1data$Condition <- ceiling(ex1data$Condition/2)

sublist <- unique(ex1data$Subject)
explist <- unique(ex1data$Experiment)
counter <- 0
datasetcounter <- 0
ntrials <- NULL
ncorrect <- NULL
subjects <- NULL
experiments <- NULL
datasets <- NULL
conditions <- NULL
pedestals <- NULL
targets <- NULL

for (s in 1:3){
  for (e in 1:3){
    datasetcounter <- datasetcounter + 1
    for (cond in 1:4){
      thisdipper <- subset(ex1data,Subject==sublist[s] & Experiment==explist[e] & Condition==cond)

      pedconts <- sort(unique(thisdipper$PedestalContrast))

      for (p in 1:length(pedconts)){
        thisthresh <- subset(thisdipper,PedestalContrast==pedconts[p])
        targconts <- sort(unique(thisthresh$TargetContrast))
        for (t in 1:length(targconts)){
          counter <- counter + 1
          thislevel <- subset(thisthresh,TargetContrast==targconts[t])
          ntrials[counter] <- nrow(thislevel)
          ncorrect[counter] <- sum(thislevel$IsCorrect)
          subjects[counter] <- s
          experiments[counter] <- e
          datasets[counter] <- datasetcounter
          conditions[counter] <- cond
          pedestals[counter] <- pedconts[p]
          targets[counter] <- targconts[t]
        }
      }
    }
  }
}

expdata <- data.frame(datasets,experiments,conditions,pedestals,targets,ntrials,ncorrect)

dataList = list(ncorrect = expdata$ncorrect, ntrials = expdata$ntrials, PedestalContrast = expdata$pedestals, TargetContrast = expdata$targets, c = expdata$conditions, s = expdata$datasets, Ndatasets = max(expdata$datasets), Ntotal = nrow(expdata))

  # Get MC sample of posterior:
  stanFit <- sampling(object=dipmodel, 
                      data = dataList, 
                      chains = nChains,
                      iter = (ceiling(numSavedSteps/nChains)*thinSteps+burnInSteps), 
                      warmup = burnInSteps, 
                      thin = thinSteps,
                      cores = getOption("mc.cores", 1L))

  save(file=paste0('local/FullMod.RData'),list=c('stanFit'))

@@

```{r plotstanmodels, include=FALSE}

@@

  load(file=paste0('local/FullMod.RData'))

  draws <- extract(stanFit)

  p <- map_estimate(draws$pmu+draws$qmu)
  q <- map_estimate(draws$qmu)
  m <- map_estimate(draws$mmu)
  k <- map_estimate(draws$kmu)
  S <- map_estimate(draws$Smu)
  Z <- map_estimate(draws$Zmu)
  w <- map_estimate(draws$wmu)
  paramvect <- c(p,q,m,S,Z,w,k)

  plist <- draws$pmu+draws$qmu
  qlist <- draws$qmu
  mlist <- draws$mmu
  klist <- draws$kmu
  Slist <- draws$Smu
  Zlist <- draws$Zmu
  wlist <- draws$wmu

  allsamples <- array(0,dim=c(7,length(plist)))
    allsamples[1,] <- plist
    allsamples[2,] <- qlist
    allsamples[3,] <- mlist
    allsamples[4,] <- Slist
    allsamples[5,] <- Zlist
    allsamples[6,] <- wlist
    allsamples[7,] <- klist

pdf('Figures/stanoutput.pdf',width=8,height=8)

layout(matrix(c(1,2,3,4,8,8,8,5,8,8,8,6,8,8,8,7),4,4,byrow=TRUE))
par(mar=c(3,2,1,1))

@@

ticklocsx <- seq(-18,24,6)
ticklabelsx <- c('1/8','1/4','1/2',1,2,4,8,16)
ticklocsy <- seq(0,1,0.2)
ticklabelsy <- ticklocsy
paramorder <- c('p','q','m','w','S','Z','k')
paramvalorder <- c(1,2,3,6,4,5,7)

@@

  plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(-18,24), ylim=c(0,1))
  axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)
  # axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
  mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.3)
  # mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)
  title(xlab=paramorder[param], col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)

    svals <- 20*log10(allsamples[paramvalorder[param],])
    a <- density(svals)
    a$y <- a$y/max(a$y)
    # lines(a$x,a$y,col=collist[expt])
polygon(a$x,a$y,border=NA,col='grey')

@@

pedlist <- -12:39
pedlevelsC <- 10^(pedlist/20)
pedlevelsC[1] <- 0
allpred <- array(0,dim=c(4,length(pedlist)))

  for (cond in 1:4){
    for (pedlev in 1:length(pedlist)){
      allpred[cond,pedlev] <- discriminate(paramvect,pedlevelsC[pedlev],cond)
    }
  }

allpreddB <- 20*log10(allpred)

par(mar=c(5,4,2,2))

@@

plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(-12,36), ylim=c(-12,39))
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.3)
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)
title(xlab="Pedestal contrast (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)
title(ylab="Threshold (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)

@@

for (cond in 1:4){
  lines(pedlist,allpreddB[cond,],col=collist[cond],lwd=2)
}
legend(-12,36,c('Monocular','Binocular','Half-bin','Dichoptic'),pch=16,col=collist,box.lwd=2,cex=1.5)

dev.off()
}

@@

  dev.off()
}

@@

load('local/MCSdata.RData')
meanslopesMCS <- 10^(apply(log10(slopematrix),c(2,3),mean))
meanslopesMCS <- round(meanslopesMCS,digits=2)

@@

titlelist <- c('(a) Achromatic target','(b) R/G target','(c) B/Y target','(d) Achromatic densities','(e) R/G densities','(f) B/Y densities','(g) Threshold elevation','(h) Psychometric slope')

    pdf(paste0('Figures/MCSdata.pdf'),width=12,height=12)

    layout(matrix(c(1,1,2,2,3,3,4,4,5,5,6,6,0,7,7,8,8,0),3,6,byrow=TRUE))

@@

outlinecolours <- c('black','black','red','cornflowerblue')
shadingcolours <- c('black','white','green','gold')

ticklocsx <- seq(-12,42,6)
ticklabelsx <- c('1/4','1/2','1','2','4','8','16','32','64','128')
ticklocsy <- seq(0,24,6)
ticklabelsy <- c('1','2','4','8','16')

thd <- 0.28

    par(pty="s")  
 plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(-12,42), ylim=c(0,24))

axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.5)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)
title(xlab="Target contrast (%)", col.lab=rgb(0,0,0), line=2, cex.lab=1.5)    # titles for axes
title(ylab=expression(paste('Weibull ', beta)), col.lab=rgb(0,0,0), line=2, cex.lab=1.5)
  title(main=titlelist[cond+3],cex.main=2)

for (block in 1:4){
condtoplot <- cond + ((block-1)*3)

xvals <- allposX[condtoplot,]
yvals <- 10.3/(allposY[condtoplot,ncol(allposY):1]/C)
zvals <- allposteriors[condtoplot,,ncol(allposY):1]

binz <- zvals/max(zvals)
iind <- which(binz>thd,arr.ind=TRUE)

xy <- matrix(0,nrow=nrow(iind),ncol=2)
xy[,1] <- xvals[iind[,1]]
xy[,2] <- yvals[iind[,2]]
hullobj <- ashape(xy,alpha=1)
polys <- extract_polygons(hullobj)
polygon(polys$x,20*log10(polys$y),col=shadingcolours[block],border=NA)
lines(polys$x[c(1:nrow(polys),1)],20*log10(polys$y[c(1:nrow(polys),1)]),lwd=2,col=outlinecolours[block])
}

for (block in 1:4){
condtoplot <- cond + ((block-1)*3)
points(MCSthresh[,condtoplot],20*log10(MCSslope[,condtoplot]),pch=21,col=outlinecolours[block],bg=shadingcolours[block])
}
}

@@

load('local/MCSdata.RData')
meanslopesMCS <- 10^(apply(log10(slopematrix),c(2,3),mean))
meanslopesMCS <- round(meanslopesMCS,digits=2)
contrastsdB <- contrastsdB[1,,]

@@

```{r dipperfig, fig.cap="Dipper functions from Experiment 1, averaged across three participants. Error bars show ±1SE across participants. Note that contrast values are expressed as a percentage of the maximum displayable contrast (see Procedures for details). Curves show...", fig.align="center", echo=FALSE}
